Fast Sorting on Modern Computers

### 2005-07-31 docs/tech_report/code/cycles_script.tex
done filename renaming inside files

\Head{}
\File{cycles\_script}{2004}{4}{29}{0:06}{773}
\L{\LB{\K{set}_\V{key}_\V{top}_\V{left}_\V{Left}_\V{reverse}}}
\L{\LB{\K{set}_\V{xrange}_[\N{4096}:\N{4194304}]}}
\L{\LB{\K{set}_\V{yrange}_[\N{0}:]}}
\L{\LB{\K{set}_\V{xtics}_(\N{8192},\N{16384},\N{32768},\N{65536},\N{131072},\N{262144},\N{524288},\N{1048576},\N{2097152})}}
\L{\LB{\K{set}_\V{format}_\V{x}_\S{}\3\%8.0f\3\SE{}}}
\L{\LB{\K{set}_\V{logscale}_\V{axis}}}
\L{\LB{\K{set}_\V{data}_\V{style}_\V{linespoints}}}
\L{\LB{\K{set}_\V{term}_\V{postscript}_\V{eps}}}
\L{\LB{\K{set}_\V{output}_\S{}{'}.\,./docs/plots/all\_cycles.eps{'}\SE{}}}
\L{\LB{\K{set}_\V{xlabel}_\S{}{'}set_size_in_keys{'}\SE{}}}
\L{\LB{\K{set}_\V{ylabel}_\S{}{'}Cycles_per_key{'}\SE{}}}
\L{\LB{\V{plot}_\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{5}_\V{title}_\S{}{'}base_quicksort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{6}_\V{title}_\S{}{'}cache_quicksort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{7}_\V{title}_\S{}{'}multi_quicksort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{10}_\V{title}_\S{}{'}base_mergesort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{14}_\V{title}_\S{}{'}double_tiled_mergesort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{16}_\V{title}_\S{}{'}double_multi\-mergesort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{9}_\V{title}_\S{}{'}base_radixsort{'}\SE{}}}

>>>>

\Head{}
\File{cycles\_script}{2004}{4}{29}{0:06}{773}
\L{\LB{\K{set}_\V{key}_\V{top}_\V{left}_\V{Left}_\V{reverse}}}
\L{\LB{\K{set}_\V{xrange}_[\N{4096}:\N{4194304}]}}
\L{\LB{\K{set}_\V{yrange}_[\N{0}:]}}
\L{\LB{\K{set}_\V{xtics}_(\N{8192},\N{16384},\N{32768},\N{65536},\N{131072},\N{262144},\N{524288},\N{1048576},\N{2097152})}}
\L{\LB{\K{set}_\V{format}_\V{x}_\S{}\3\%8.0f\3\SE{}}}
\L{\LB{\K{set}_\V{logscale}_\V{axis}}}
\L{\LB{\K{set}_\V{data}_\V{style}_\V{linespoints}}}
\L{\LB{\K{set}_\V{term}_\V{postscript}_\V{eps}}}
\L{\LB{\K{set}_\V{output}_\S{}{'}.\,./docs/plots/all\_cycles.eps{'}\SE{}}}
\L{\LB{\K{set}_\V{xlabel}_\S{}{'}set_size_in_keys{'}\SE{}}}
\L{\LB{\K{set}_\V{ylabel}_\S{}{'}Cycles_per_key{'}\SE{}}}
\L{\LB{\V{plot}_\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{5}_\V{title}_\S{}{'}base_quicksort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{6}_\V{title}_\S{}{'}memory_tuned_quicksort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{7}_\V{title}_\S{}{'}multi_quicksort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{10}_\V{title}_\S{}{'}base_mergesort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{14}_\V{title}_\S{}{'}double_aligned_mergesort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{16}_\V{title}_\S{}{'}double_multi\-mergesort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{9}_\V{title}_\S{}{'}base_radixsort{'}\SE{}}}

### 2005-07-31 docs/tech_report/merge.tex
added comparisons with lamarca to the conclusions
fixing up lots of things
done filename renaming inside files

\section{Predictor results}

\sixplot{algorithm_n_0.eps}{algorithm_s_0.eps}{base_mergesort_0.eps}{double_tiled_mergesort_0.eps}{double_multi_mergesort_0.eps}{double_multi_mergesort_0.eps}{mergesorts}

Figure \ref{mergesorts-1} shows that algorithm N has the same number of branches
and branch mispredictions as algorithm S, shown in \ref{mergesorts-2}. Base
mergesort has the same number of mispredictions (see Figure \ref{mergesorts-3},
indicating a possible minimum number. It has a lower number of branches,
probably due to performing a presort first. The presort has approximately 6.8
misses per key (not shown in the graphs). In later mergesorts, an insertion sort
is used instead.
%TODO ohw much better does this do? if not, move to other section

@@

\section{Future work}
Several avenues could be pursued in order to improve the algorithms presented
here. In multi-mergesort, LaMarca spoke of tagging keys as they went into the
cache. Here, a sorted list was used in its place. Implementing and investigating
tagging and comparing of the results of the two techniques would be interesting.

>>>>

\section{Predictor Results}

\sixplot{algorithm_n_0.eps}{algorithm_s_0.eps}{base_mergesort_0.eps}{double_aligned_mergesort_0.eps}{double_aligned_multi_mergsort_0.eps}{double_aligned_multi_mergsort_0.eps}{mergesorts}

Figure \ref{mergesorts-1} shows that algorithm N has the same number of branches
and branch mispredictions as algorithm S, shown in \ref{mergesorts-2}. Base
mergesort has the same number of mispredictions (see Figure \ref{mergesorts-3},
indicating a possible minimum number. It has a lower number of branches,
probably due to performing a presort first. The presort has approximately 6.8
misses per key (not shown in the graphs). In later mergesorts, an insertion sort
is used instead.

@@

\section{Future Work}
Several avenues could be pursued in order to improve the algorithms presented
here. In multi-mergesort, LaMarca spoke of tagging keys as they went into the
cache. Here, a sorted list was used in its place. Implementing and investigating
tagging and comparing of the results of the two techniques would be interesting.

### 2005-07-31 docs/tech_report/quick.tex
minor edit
fixing up lots of things
done filename renaming inside files

\section{Predictor results}

%TODO a bit to rewrite
\subsection{Different sorts}

\sixplot{base_quicksort_0.eps}{base_quicksort9_0.eps}{cache_quicksort_0.eps}{multi_quicksort_0.eps}{sequential_multi_quicksort_0.eps}{}{base, cache and multi quicksorts and base quicksort9}

Firstly, cache quicksort and base quicksort only idffer in how they deal with
the insertion. Base quicksort and base quicksort9 have similar numbers of
insertion branches, and a very similar ratio (2/1 correct/incorrect) of
mispredictions. Comparing base quicksort to cahce quicksort (they both use
median of 3), there is actually a decrease in the number of branches handled by
insertion sort, instead of an increase. An increase is expected (thats why
sedgewick moves the insertion to the end), though when you take into account
the sentinel and the outer loop, there probably is an increase in instructions
at least. Base quicksort has 35\% miss rate, but cache has a 34\%. Not a big
difference, but the point is that there is a reduction in branches, but of the
reduced branches, 50\% were unpredictable. I dont think theres anything
interesting in this fact, however.

@@

%TO here

\subsection{Different medians}

>>>>

\section{Predictor Results}

%TODO a bit to rewrite
\subsection{Different Sorts}

\sixplot{base_quicksort_0.eps}{base_quicksort9_0.eps}{memory_tuned_quicksort_0.eps}{multi_quicksort_0.eps}{sequential_multi_quicksort_0.eps}{}{base, cache and multi quicksorts and base quicksort9}

Firstly, cache quicksort and base quicksort only differ in how they deal with
the insertion. Base quicksort and base quicksort9 have similar numbers of
insertion branches, and a very similar ratio (2/1 correct/incorrect) of
mispredictions. Comparing base quicksort to cahce quicksort (they both use
median of 3), there is actually a decrease in the number of branches handled by
insertion sort, instead of an increase. An increase is expected (thats why
sedgewick moves the insertion to the end), though when you take into account
the sentinel and the outer loop, there probably is an increase in instructions
at least. Base quicksort has 35\% miss rate, but cache has a 34\%. Not a big
difference, but the point is that there is a reduction in branches, but of the
reduced branches, 50\% were unpredictable. I dont think theres anything
interesting in this fact, however.

@@

%TODO to  here

\subsection{Different Medians}

### 2005-07-31 docs/tech_report/result_list.tex
fixing up lots of things
done filename renaming inside files

\label{result_list}
\afterpage{
\smallplot{on2_insertsort}{insertion sort}{}{}{}{}{}
\smallplot{on2_selectsort}{selection sort}{}{}{}{}{}
\smallplot{on2_bubblesort}{bubblesort}{}{}{}{}{}
\smallplot{on2_bubblesort2}{improved bubblesort}{}{}{}{}{}
\smallplot{on2_shakersort}{shakersort}{}{}{}{}{}
\smallplot{on2_shakersort2}{improved shakersort}{}{}{}{}{}
%
\smallplot{base_heapsort}{base heapsort}{}{}{}{}{}
\smallplot{cache4_heapsort}{cache heapsort (4-heap)}{}{}{}{}{}
\smallplot{cache_heapsort}{cache heapsort (8-heap)}{}{}{}{}{}
%
\smallplot{algorithm_n}{algorithm N}{}{}{}{}{}
\smallplot{algorithm_s}{algorithm S}{}{}{}{}{}
\smallplot{base_mergesort}{base mergesort}{}{}{}{}{}
\smallplot{tiled_mergesort}{tiled mergesort}{}{}{}{}{}
\smallplot{multi_mergesort}{multi-mergesort}{}{}{}{}{}
\smallplot{double_tiled_mergesort}{double tiled mergesort}{}{}{}{}{}
\smallplot{double_multi_mergesort}{double multi-mergesort}{}{}{}{}{}
%
\smallplot{base_quicksort1}{base quicksort (no median)}{}{}{}{}{}
\smallplot{base_quicksort}{base quicksort (median-of-3)}{}{}{}{}{}
\smallplot{base_quicksort5}{base quicksort (pseudo-median-of-5)}{}{}{}{}{}
\smallplot{base_quicksort7}{base quicksort (pseudo-median-of-7)}{}{}{}{}{}
\smallplot{base_quicksort9}{base quicksort (pseudo-median-of-9)}{}{}{}{}{}
\smallplot{cache_quicksort}{memory-tuned quicksort}{}{}{}{}{}
\smallplot{multi_quicksort}{multi-quicksort (binary search)}{}{}{}{}{}
\smallplot{multi_quicksort_seq}{multi-quicksort (sequential search)}{}{}{}{}{}
%
\smallplot{base_radixsort}{base radixsort}{}{}{}{}{}
\smallplot{cache_radixsort}{cache radixsort}{}{}{}{}{}
\smallplot{aligned_cache_radixsort}{aligned cache radixsort}{}{}{}{}{}
%
\smallplot{shellsort}{shellsort}{}{}{}{}{}
\smallplot{improved_shellsort}{improved shellsort}{}{}{}{}{}
}

>>>>

\label{result_list}
\afterpage{
\smallplot{insertsort}{insertion sort}{}{}{}{}{}
\smallplot{selectsort}{selection sort}{}{}{}{}{}
\smallplot{bubblesort}{bubblesort}{}{}{}{}{}
\smallplot{bubblesort2}{improved bubblesort}{}{}{}{}{}
\smallplot{shakersort}{shakersort}{}{}{}{}{}
\smallplot{shakersort2}{improved shakersort}{}{}{}{}{}
%
\smallplot{base_heapsort}{base heapsort}{}{}{}{}{}
\smallplot{memory_tuned4_heapsort}{cache heapsort (4-heap)}{}{}{}{}{}
\smallplot{memory_tuned_heapsort}{cache heapsort (8-heap)}{}{}{}{}{}
%
\smallplot{algorithm_n}{algorithm n}{}{}{}{}{}
\smallplot{algorithm_s}{algorithm s}{}{}{}{}{}
\smallplot{base_mergesort}{base mergesort}{}{}{}{}{}
\smallplot{tiled_mergesort}{tiled mergesort}{}{}{}{}{}
\smallplot{multi_mergesort}{multi-mergesort}{}{}{}{}{}
\smallplot{double_aligned_mergesort}{double tiled mergesort}{}{}{}{}{}
\smallplot{double_aligned_multi_mergsort}{double multi-mergesort}{}{}{}{}{}
%
\smallplot{base_quicksort1}{base quicksort (no median)}{}{}{}{}{}
\smallplot{base_quicksort}{base quicksort (median-of-3)}{}{}{}{}{}
\smallplot{base_quicksort5}{base quicksort (pseudo-median-of-5)}{}{}{}{}{}
\smallplot{base_quicksort7}{base quicksort (pseudo-median-of-7)}{}{}{}{}{}
\smallplot{base_quicksort9}{base quicksort (pseudo-median-of-9)}{}{}{}{}{}
\smallplot{memory_tuned_quicksort}{memory-tuned quicksort}{}{}{}{}{}
\smallplot{multi_quicksort}{multi-quicksort (binary search)}{}{}{}{}{}
\smallplot{multi_quicksort_seq}{multi-quicksort (sequential search)}{}{}{}{}{}
%
\smallplot{base_radixsort}{base radixsort}{}{}{}{}{}
\smallplot{memory_tuned_radixsort}{cache radixsort}{}{}{}{}{}
\smallplot{aligned_memory_tuned_radixsort}{aligned cache radixsort}{}{}{}{}{}
%
\smallplot{shellsort}{shellsort}{}{}{}{}{}
\smallplot{improved_shellsort}{improved shellsort}{}{}{}{}{}
}

### 2005-07-31 docs/tech_report/software_predictor.tex
removed word gonnet from document
done filename renaming inside files

\afterpage{
\sixplot{base_heapsort_0.eps}{cache4_heapsort_0.eps}{cache4_heapsort_1.eps}{cache_heapsort_0.eps}{cache_heapsort_1.eps}{}{base and cache heapsorts}

\sixplot{base_quicksort_0.eps}{base_quicksort9_0.eps}{cache_quicksort_0.eps}{multi_quicksort_0.eps}{sequential_multi_quicksort_0.eps}{}{base, cache and multi quicksorts and base quicksort9}
\sixplot{base_quicksort1_0.eps}{base_quicksort_0.eps}{base_quicksort5_0.eps}{base_quicksort7_0.eps}{base_quicksort9_0.eps}{}{base quicksort without a median, with median of 3, and with pseudo median of 5, 7 and 9}

\sixplot{algorithm_n_0.eps}{algorithm_n_1.eps}{algorithm_s_0.eps}{algorithm_s_1.eps}{}{}{Knuth's Mergesorts}
\sixplot{base_mergesort_0.eps}{base_mergesort_1.eps}{base_mergesort_2.eps}{base_mergesort_3.eps}{base_mergesort_4.eps}{base_mergesort_5.eps}{base mergesort}
\sixplot{double_tiled_mergesort_0.eps}{double_tiled_mergesort_1.eps}{double_tiled_mergesort_2.eps}{double_tiled_mergesort_3.eps}{double_tiled_mergesort_4.eps}{double_tiled_mergesort_5.eps}{double tiled mergesort - part 1}
\sixplot{double_tiled_mergesort_6.eps}{double_tiled_mergesort_7.eps}{double_multi_mergesort_0.eps}{double_multi_mergesort_1.eps}{double_multi_mergesort_2.eps}{double_multi_mergesort_3.eps}{double tiled mergesort - part 2, and double multi mergesort - part 1}
\sixplot{double_multi_mergesort_4.eps}{double_multi_mergesort_5.eps}{double_multi_mergesort_6.eps}{double_multi_mergesort_7.eps}{double_multi_mergesort_8.eps}{double_multi_mergesort_9.eps}{double multi-mergesort - part 2}

\sixplot{gonnet_shellsort_0.eps}{gonnet_improved_shellsort_0.eps}{}{}{}{}{Shellsorts}
}

>>>>

\afterpage{
\sixplot{base_heapsort_0.eps}{memory_tuned4_heapsort_0.eps}{memory_tuned4_heapsort_1.eps}{memory_tuned_heapsort_0.eps}{memory_tuned_heapsort_1.eps}{}{base and cache heapsorts}

\sixplot{base_quicksort_0.eps}{base_quicksort9_0.eps}{memory_tuned_quicksort_0.eps}{multi_quicksort_0.eps}{sequential_multi_quicksort_0.eps}{}{base, cache and multi quicksorts and base quicksort9}
\sixplot{base_quicksort1_0.eps}{base_quicksort_0.eps}{base_quicksort5_0.eps}{base_quicksort7_0.eps}{base_quicksort9_0.eps}{}{base quicksort without a median, with median of 3, and with pseudo median of 5, 7 and 9}

\sixplot{algorithm_n_0.eps}{algorithm_n_1.eps}{algorithm_s_0.eps}{algorithm_s_1.eps}{}{}{Knuth's Mergesorts}
\sixplot{base_mergesort_0.eps}{base_mergesort_1.eps}{base_mergesort_2.eps}{base_mergesort_3.eps}{base_mergesort_4.eps}{base_mergesort_5.eps}{base mergesort}
\sixplot{double_aligned_mergesort_0.eps}{double_aligned_mergesort_1.eps}{double_aligned_mergesort_2.eps}{double_aligned_mergesort_3.eps}{double_aligned_mergesort_4.eps}{double_aligned_mergesort_5.eps}{double tiled mergesort - part 1}
\sixplot{double_aligned_mergesort_6.eps}{double_aligned_mergesort_7.eps}{double_aligned_multi_mergsort_0.eps}{double_aligned_multi_mergsort_1.eps}{double_aligned_multi_mergsort_2.eps}{double_aligned_multi_mergsort_3.eps}{double tiled mergesort - part 2, and double multi mergesort - part 1}
\sixplot{double_aligned_multi_mergsort_4.eps}{double_aligned_multi_mergsort_5.eps}{double_aligned_multi_mergsort_6.eps}{double_aligned_multi_mergsort_7.eps}{double_aligned_multi_mergsort_8.eps}{double_aligned_multi_mergsort_9.eps}{double multi-mergesort - part 2}

\sixplot{shellsort_0.eps}{improved_shellsort_0.eps}{}{}{}{}{Shellsorts}
}
